% modified from NeurIPS 2022 Latex template
% authors: Yixin Zhu (yixin.zhu@pku.edu.cn), Liangru Xiang
% v1.0: 2022.07.21

\documentclass{article}
\PassOptionsToPackage{numbers,compress}{natbib}
\usepackage[final]{template22}
\input{config.tex}

\title{Block Stacking with Robot Motion Planners}

\author{%
  Your Name \\
  Undergraduate Student, Department of Computer Science\\
  Your University\\
  \texttt{yourName@xxx.edu.cn} \\
}

\begin{document}
\maketitle

\begin{abstract}
Block stacking is a classic manipulation task that requires a robot to pick up a set of cubes and place them one on top of another to form a stable tower.
Solving this task cleanly exposes the core challenges of robotic manipulation: object pose estimation, grasp planning, collision-free motion planning, and reactive control.
In this report, I study how to implement a block-stacking pipeline using a robot arm equipped with a parallel-jaw gripper under progressively relaxed assumptions.
I first present a structured pseudo-code solution that relies on off-the-shelf motion planners and discuss the arguments needed at each API call.
I then address a practical collision issue that arises when planning directly to a grasp pose, and propose a simple pre-grasp waypoint strategy to resolve it.
Finally, I discuss how to extend the system to an unconstrained setting where cube poses are unknown and must be estimated online with a depth camera, identifying the essential perception, planning, and control modules required for robust task completion.
\end{abstract}


\section{Introduction}
\label{sec:introduction}

Block stacking is a foundational benchmark in robot manipulation research.
Given $N$ cubes placed at known (or unknown) positions on a table, the robot must pick them up in order and stack them to form a tower of height $N$.
Despite its apparent simplicity, the task exercises the full manipulation pipeline: workspace sensing, grasp planning, collision-free motion planning, and execution feedback.

In the following sections, I address three questions.
\Cref{sec:pseudocode} presents a complete pseudo-code solution assuming cube poses are known.
\Cref{sec:collision} proposes a simple strategy to avoid end-effector collisions during cube grasping.
\Cref{sec:depth} discusses the additional modules needed when cube poses are unknown and a depth camera is the primary sensor.

\section{Pseudo code for block stacking}
\label{sec:pseudocode}

I assume the robot has a 6-DoF arm with a parallel-jaw gripper, that the 6-DoF pose of every cube is known in the robot base frame, and that the arm's forward/inverse kinematics are available.
The core API is \texttt{motion\_plan}, whose signature and semantics are described below; all other helper functions are thin wrappers around standard robot-middleware calls.

\paragraph{Key API: \texttt{motion\_plan}}
\begin{verbatim}
trajectory = motion_plan(
    start,            # current joint configuration q_start (or "current")
    goal,             # target end-effector pose T_goal (SE(3)) or joint config
    collision_scene,  # list of collision objects currently in the workspace
    motion_type,      # FREE (joint-space RRT/PRM) | CARTESIAN (line in task space)
    constraints       # optional: orientation lock, joint limits, etc.
)
\end{verbatim}

\noindent\textbf{Arguments.}
\textit{start} and \textit{goal} define the boundary conditions of the plan.
\textit{goal} is usually expressed as a target end-effector pose $T \in SE(3)$ that is first solved for joint angles via inverse kinematics; alternatively a joint configuration $\mathbf{q}$ can be passed directly.
\textit{collision\_scene} is a list of axis-aligned bounding boxes (or mesh primitives) for every rigid body in the scene---table, cubes not yet grasped, cubes already stacked.
\textit{motion\_type} selects the planner: \texttt{FREE} uses a sampling-based planner (\eg, RRT-Connect) for unconstrained free-space motion; \texttt{CARTESIAN} generates a straight-line interpolated path in task space for approach and retreat strokes.
\textit{constraints} may impose, \eg, a fixed wrist orientation to keep the gripper level while carrying a cube.

\paragraph{Helper functions.}

\begin{itemize}[leftmargin=*,noitemsep,topsep=0pt]
  \item \texttt{get\_object\_pose(obj)} $\to T_\text{obj} \in SE(3)$: returns the 6-DoF pose of \texttt{obj} in the robot base frame.
  \item \texttt{compute\_grasp\_pose(T\_cube, gripper\_depth)} $\to T_\text{grasp}$: offsets the cube centre upward by half the cube height and back by \texttt{gripper\_depth} along the approach axis; adds a fixed wrist orientation (gripper fingers horizontal, approaching from above).
  \item \texttt{compute\_pre\_grasp\_pose(T\_grasp, approach\_offset)} $\to T_\text{pre}$: translates $T_\text{grasp}$ by \texttt{approach\_offset} ($\sim$\,10--15\,cm) along the gripper approach axis (\ie, straight up for a top-down grasp).
  \item \texttt{compute\_pre\_place\_pose(T\_place, approach\_offset)} $\to T_\text{pre\_place}$: analogous to \texttt{compute\_pre\_grasp\_pose} but applied to the placement pose; translates $T_\text{place}$ upward by \texttt{approach\_offset} to produce a safe above-stack waypoint before lowering.
  \item \texttt{compute\_place\_pose(T\_stack\_base, stack\_height, cube\_size)} $\to T_\text{place}$: adds $\texttt{stack\_height} \times \texttt{cube\_size}$ to the $z$-component of the stack base pose.
  \item \texttt{gripper\_close()} / \texttt{gripper\_open()}: command the gripper to close/open.
  \item \texttt{update\_collision\_scene(obj, new\_pose)}: informs the motion planner that \texttt{obj} has moved to \texttt{new\_pose} (or has been attached to / detached from the gripper).
\end{itemize}

\paragraph{Pseudo code.}

\Cref{alg:block_stacking} gives the complete procedure for stacking $N$ cubes.

\begin{algorithm}[t]
\caption{Block stacking with a robot arm}
\label{alg:block_stacking}
\begin{algorithmic}[1]
\REQUIRE cube list $\mathcal{C} = [c_1, \ldots, c_N]$, stack base pose $T_\text{base}$, cube size $s$
\FOR{$i = 1$ \TO $N$}
  \STATE $T_\text{cube} \leftarrow$ \texttt{get\_object\_pose}$(c_i)$
  \STATE $T_\text{grasp} \leftarrow$ \texttt{compute\_grasp\_pose}$(T_\text{cube},\ \textit{gripper\_depth})$
  \STATE $T_\text{pre} \leftarrow$ \texttt{compute\_pre\_grasp\_pose}$(T_\text{grasp},\ \textit{approach\_offset})$
  \STATE \COMMENT{Move to pre-grasp waypoint (free-space planning, avoid all objects)}
  \STATE \texttt{execute}(\texttt{motion\_plan}(\texttt{``current''}, $T_\text{pre}$, $\mathcal{S}$, \texttt{FREE}))
  \STATE \COMMENT{Straight-line descent to grasp pose (Cartesian, no cube in collision set)}
  \STATE \texttt{execute}(\texttt{motion\_plan}($T_\text{pre}$, $T_\text{grasp}$, $\mathcal{S} \setminus \{c_i\}$, \texttt{CARTESIAN}))
  \STATE \texttt{gripper\_close}()
  \STATE \texttt{update\_collision\_scene}$(c_i,\ \text{attach to gripper})$
  \STATE \COMMENT{Lift straight up back to pre-grasp height}
  \STATE \texttt{execute}(\texttt{motion\_plan}($T_\text{grasp}$, $T_\text{pre}$, $\mathcal{S}$, \texttt{CARTESIAN}))
  \STATE $T_\text{place} \leftarrow$ \texttt{compute\_place\_pose}$(T_\text{base},\ i-1,\ s)$
  \STATE $T_\text{pre\_place} \leftarrow$ \texttt{compute\_pre\_place\_pose}$(T_\text{place},\ \textit{approach\_offset})$
  \STATE \COMMENT{Carry cube to pre-place waypoint (free-space, orientation locked)}
  \STATE \texttt{execute}(\texttt{motion\_plan}($T_\text{pre}$, $T_\text{pre\_place}$, $\mathcal{S}$, \texttt{FREE}, \textit{orientation\_lock}))
  \STATE \COMMENT{Lower straight down to place pose}
  \STATE \texttt{execute}(\texttt{motion\_plan}($T_\text{pre\_place}$, $T_\text{place}$, $\mathcal{S}$, \texttt{CARTESIAN}))
  \STATE \texttt{gripper\_open}()
  \STATE \texttt{update\_collision\_scene}$(c_i,\ T_\text{place})$
  \STATE $\mathcal{S} \leftarrow \mathcal{S} \cup \{c_i\ \text{at}\ T_\text{place}\}$
  \STATE \COMMENT{Retreat straight upward before next pick}
  \STATE \texttt{execute}(\texttt{motion\_plan}($T_\text{place}$, $T_\text{pre\_place}$, $\mathcal{S}$, \texttt{CARTESIAN}))
\ENDFOR
\end{algorithmic}
\end{algorithm}

\paragraph{Argument computation.}
The grasp pose $T_\text{grasp}$ is computed so that the gripper centre aligns with the cube centre in $x$ and $y$, is raised by $s/2$ in $z$ (top-face contact), and approaches from directly above (rotation: gripper $z$-axis pointing downward).
The pre-grasp offset is chosen large enough to clear the cube geometry---typically $0.10$--$0.15$\,m along the approach axis.
The place pose $z$-offset is $(i-1) \cdot s$, stacking cubes flush with no gap.
The collision scene $\mathcal{S}$ is updated after every gripper action so subsequent plans respect the new configuration; when a cube is attached to the gripper it is removed from the free-object list and treated as part of the robot body.


\section{Avoiding end-effector collision when grasping}
\label{sec:collision}

Planning directly from the robot's current pose to the grasp pose $T_\text{grasp}$ allows the sampling-based planner to generate paths that approach the cube from the side or from below, causing the gripper fingers to collide with the cube body before closing.

\paragraph{Pre-grasp waypoint strategy.}
The simplest and most effective remedy is to insert a \emph{pre-grasp waypoint} $T_\text{pre}$, obtained by translating $T_\text{grasp}$ by a fixed offset $d$ ($\sim$\,10--15\,cm) along the gripper's approach axis (straight up for a top-down grasp, \cref{alg:block_stacking} lines 3--6).
The motion from the current pose to $T_\text{pre}$ is planned freely in joint space---the cube is still in the collision scene, so the planner avoids it naturally.
The final descent from $T_\text{pre}$ to $T_\text{grasp}$ is executed as a \emph{straight-line Cartesian motion} with the cube temporarily removed from the collision set (because the gripper is intentionally entering the cube's bounding volume to close around it).
This guarantees a clean top-down approach with no lateral sweeping near the cube surface.
After grasping, the inverse retreat motion lifts the cube straight up before free-space replanning, preventing collisions with neighbouring objects.


\section{Open question: block stacking with a depth camera}
\label{sec:depth}

When the simplicity assumptions (known cube poses, clean workspace) cannot be guaranteed, a depth camera mounted on the robot's wrist or on a fixed overhead rig provides the raw 3D data needed to close the perception--action loop.
The following essential modules and procedures are required.

\paragraph{Essential modules.}

\begin{enumerate}[leftmargin=*,noitemsep,topsep=0pt]
  \item \textbf{Depth-image acquisition and preprocessing.}
        The depth camera streams aligned RGB-D frames.
        A preprocessing pipeline filters noise (bilateral or guided filtering), removes background via a depth threshold, and converts the depth map to a registered 3D point cloud in the robot base frame using the known camera intrinsics and the camera-to-base extrinsic calibration (obtained once offline via a checkerboard).

  \item \textbf{3D scene segmentation.}
        The point cloud is segmented into individual object clusters using, \eg, plane fitting (RANSAC) to remove the table, followed by Euclidean cluster extraction or instance-segmentation-based methods.
        Each cluster is associated with a cube hypothesis.

  \item \textbf{6-DoF object pose estimation.}
        For each cube cluster, a pose estimation module fits a known CAD model (a cube of known side length $s$) to the cluster, yielding a 6-DoF pose $T_\text{cube}$.
        Classical approaches use iterative closest point (ICP); learning-based approaches (\eg, FoundationPose, FFB6D) are more robust to partial occlusion.

  \item \textbf{Collision map construction.}
        The full point cloud (excluding the detected cubes) is voxelised into an occupancy grid or OctoMap and fed to the motion planner as the collision scene, enabling obstacle-aware path planning without manual scene description.

  \item \textbf{Grasp planner.}
        Given the estimated cube pose, a grasp planner (\eg, GraspNet or an analytical antipodal sampler) generates candidate grasp poses ranked by stability and reachability.
        The top-ranked reachable grasp is passed to the motion planner.

  \item \textbf{Collision-free motion planner.}
        As described in \cref{sec:pseudocode}, a sampling-based planner (\eg, RRT-Connect via MoveIt) generates joint-space trajectories that avoid the live collision map.

  \item \textbf{Execution monitor and recovery.}
        Force/torque and depth feedback are monitored during execution.
        If the grasp fails (detected by gripper width after closing), the system re-observes the scene and replans.
        If a cube is displaced mid-execution, the perception module detects the change and the task planner re-sequences the remaining picks.
\end{enumerate}

\paragraph{Procedure.}
\begin{enumerate}[leftmargin=*,noitemsep,topsep=0pt]
  \item Observe the workspace: capture one or more RGB-D frames; optionally move the camera to reduce occlusions.
  \item Preprocess and segment the point cloud; estimate 6-DoF poses of all cubes.
  \item Build the collision map from the full scene point cloud.
  \item Select the next cube to grasp (typically the topmost unobstructed cube); compute a grasp pose and a pre-grasp waypoint.
  \item Plan and execute the pick-and-place sequence from \cref{alg:block_stacking}, using the live collision map.
  \item After placing, re-observe the stack to verify placement accuracy; if the placed cube has slipped, correct before the next pick.
  \item Repeat until all cubes are stacked.
\end{enumerate}

\noindent The depth camera thus replaces the assumption of known poses with an online perception pipeline, while the motion planning and execution layers remain largely unchanged.

\bibliographystyle{plainnat}
\bibliography{reference}

\appendix

\section{Appendix}

No additional material.

\end{document}